---
title: "Initial Korean covid 19"
author: "Mairani"
date: "5/30/24"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female  
77 of the data entries were coded sex  as blank, so we omit these.  
state of a patient must be died or alive - we   adjust that as well     
  
# Your Question - Is dying independent of sex?
$H_0$:
There is no difference in the death rates between men and women in korea due to covid.  

$H_A$:There is a difference in the death rates between men and women in korea due to covid.  





# Methods

result is a categorical variable, sex is   categorical  result depends on sex, we use helper   file cat~cat  

# results

## Graphical results - use ggplot commands to do this  


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
   
   This bar chart shows that more Korean men died than women because of Covid19.     

  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The percentage that were male was 64%, almost twice as many as female deaths.It shows that men were impacted at a higher rate.

##Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```




## Conclusions

